---
title: "Assignment 1"
author: "Zane Lesley"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
setwd("C:/Users/Zanel/OneDrive/Desktop/FALL 2024 STATS/FALL224753lesl0009/Lab/Lab1")
ddt=read.csv("DDT.csv")
head(ddt)
```
# Tasks

## Q1

In Class Quizzes - 10%
Labs = 10%
Chapter Quizzes - 5%
Project 1 - 3.33%
Project 2 - 6.67%
Exam 1 - 10%
Exam 2 - 10%
Assignments - 15%
Final Exam - 30%

Normal grading, but B is extended to 70%+ and C is 60%+.

## Q2

```{r}
m = with(ddt, as.numeric(factor(MILE)))
length(unique(m))

with(ddt, coplot(LENGTH~WEIGHT | RIVER*SPECIES, row=1, col=m))
```

### b)
The catfish were very similar within Length and weight, being found in different rivers as well at different miles.

### c)
factor() essentially converts the MILE into a list of values, which as.numeric then converts those unique values
into numeric values.

### d)
unique(m) makes each of those numeric values unique, consisting of only one of each type, which length gives how many of those values there
are.

### e)
No Large mouth bass or small buffalo fish exist in those rivers.

### f)
```{r}
mean(ddt[ddt$RIVER == "FCM" & ddt$SPECIES == "CCATFISH"]$DDT)
```

## Q3)
a. Quantitative 
b. Quantitative
c. Qualitative
d. Quantitative
e. Qualitative
f. Quantitative
g. Qualitative

## Q4)

simple random - select x amount of data points using a random number generator to assign them to the sample.

stratified random sampling - data points are chosen when you can categorical sort the data points into groups that are more 
similar within their strata vs across the sample's strata. We then get random samples from within those strata. 
(speed limit example 25mph, 40mph, or 55mph depending on max street speed limit)

cluster sampling - sample a certain amount from the whole population, so pick 10 of 150 programs, then examine each program for errors.

systematic sampling - select ever nth data point in the population.

## Q5)
```{r}
setwd("C:/Users/Zanel/OneDrive/Desktop/FALL 2024 STATS/FALL224753lesl0009/Assignment/ASS1")
mtbe = read.csv("MTBE.csv")
ind = sample(1:223, 5, replace=FALSE)
mtbe[ind,]
```

### a)
```{r}
mtbeo = na.omit(mtbe)
sd(mtbeo[mtbeo$Aquifier == "Bedrock",]$Depth)
```

## Q6)
```{r}
setwd("C:/Users/Zanel/OneDrive/Desktop/FALL 2024 STATS/FALL224753lesl0009/Assignment/ASS1")
eq = read.csv("EARTHQUAKE.csv")
rm(ind)
ind = sample(1:2929, 30, replace=FALSE)
eq[ind,]
```

### ai)
```{r}
plot(ts(eq$MAG))
```

### aii
```{r}
median(eq$MAGNITUDE)
```

## Q7)

Stratified Random Sampling 

The population is the fish in the rivers, the sample would be the fish they caught.

River, Species.

## Q8)
a. Bar 
b. Types of Robotic Limbs
c. Legs Only
d. 15,8,63,20
### e)
```{r}
freq = c(15, 8, 63, 20)
RL = c("None", "Both", "Legs","Wheels")
l = rep(RL, freq)
source("pareto.r")
pareto(l)
```

## Q9)
```{r}
freq = c(6, 8, 22, 3, 11)
RL = c("DDOS", "ID", "RCE","Spoofing", "PE")
l = rep(RL, freq)

pie(freq, labels=RL)
pareto(l)
```

## Q10)
```{r}
swd = read.csv("SWDEFECTS.csv", header=TRUE)
library(plotrix)
tab=table(swd$defect)
rtab=tab/sum(tab)
round(rtab, 2)
pie3D(rtab, labels=list("OK", "Defective"), main="pie plot of SWD")
```

The piechart shows that most of the software is good and not defective, while a small portion of it is still defective.

## Q11)











































