---
title: "Lab2"
author: "Zane Lesley"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

### setwd()
```{r}
setwd("C:/Users/Zanel/OneDrive/Desktop/FALL 2024 STATS/FALL224753lesl0009/Lab/Lab 2")
```

## Task 2 

### CSV
```{r}
ddt <- read.csv("EPAGAS.CSV")
head(ddt, 6)
```

## Task 3

```{r}
mpg <- ddt$MPG
```

### Z Transformation and Verification
```{r}
z = (mpg - mean(mpg)) / sd(mpg)

mean(z)
sd(z)^2
```

### Outliers
```{r}
mpg[abs(z)>=2 & abs(z)<=3]

mpg[abs(z)>3]
```

### lattice
```{r}
library(lattice)
dotplot(mpg, col= 
          ifelse(abs(z)>3, "Red",
               ifelse(abs(z)>=2 & abs(z)<=3, "Blue", "Black")
               )
        )
```

## Task 4
### Boxplot
```{r}
library(ggplot2)
ggplot(ddt, aes(y=mpg)) + geom_boxplot(fill='Black', notch=TRUE) + labs(y="Miles Per gallon", x="Z-Score") + ggtitle("Z-Score vs. Miles per Gallon") + coord_flip()
```

## Chebyshev
3/4ths of the data should lie within two standard deviations of the mean.

## Exact Porportion within 2 SD of the mean
```{r}
length(mpg[abs(z) < 2]) / length(mpg)
```

### Does Chebyshev Agree?
Yes, it agrees, we had 75% or more in our data lie within the SD of 2.

### Empirical Rule  
95% of the data should lie within two standard deviations of the mean.

### Does it correspond?
It corresponds very closely being that we got 96% of our data is within 2 SD of the mean

### Is it valid in this case?
For it to be valid, the graph of the data (mpg in this case) must be in a bell curve. We can verify this with a histogram with sd as our x and Frequency as our y.
```{r}
ggplot(ddt, aes(x = z)) + 
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(x = "z score", y = "Frequency")
```

You can see if we start grouping the close z-scores together, we get a nice bell curve which means it is valid.




















