---
title: "Lab1"
author: "Zane Lesley"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

```{r}
getwd()
```

## Task 2

```{r}
ddt <- read.csv("DDT.csv")
head(ddt, 6)
```

## Task 3

### Qualitative Variables

```{r}
names(ddt[,c("RIVER", "SPECIES")])
```

### Quantative Variables

```{r}
names(ddt[,c("MILE", "LENGTH", "WEIGHT", "DDT")])
```

### Species in DDT

```{r}
length(ddt[,"SPECIES"])
```

### Species == LMBASS, Weight > 800gms

```{r}
with(ddt, ddt[SPECIES=="LMBASS" & WEIGHT > 800,])
```

### River == SCM, DDT > 4.0
```{r}
with(ddt, ddt[RIVER == "SCM" & DDT> 4.0,])
```

## Task 4
### Mean of length of fish
```{r}
mean(ddt$LENGTH)
```

### Standard Devation of Weight
```{r}
sd(ddt$WEIGHT)
```

### Length vs Weight
```{r}
filtered_ddt <- ddt[,c("LENGTH", "WEIGHT")]
with(filtered_ddt, filtered_ddt[order(WEIGHT),])
```
No it is swapped, graph is WEIGHT vs LENGTH

### last value of v/20
```{r}
v <- 1:20
v[20] / 20
```
1

## Task 5

### Table of Rivers
```{r}
river_table = with(ddt, table(RIVER))
river_table
```

### Barplot of rivers
```{r}
library(ggplot2)
barplot(river_table, col=rainbow(4))
```
### 2D Table
```{r}
table_riv_species = with(ddt, table(RIVER, SPECIES))
table_riv_species
```

### 2D Bar Plot
```{r}
barplot(table_riv_species)
```

## Task 6
### Pie chart of Fish Species
```{r}
pie(with(ddt, table(SPECIES)))
```

### Pie Chart of Rivers
```{r}
pie(with(ddt, table(RIVER)))
```

## Task 7
### Boxplot of DDT
```{r}
with(ddt, boxplot(DDT))
```

### Boxplot of WEIGHT
```{r}
with(ddt, boxplot(WEIGHT))
```

### Boxplot of LENGTH
```{r}
with(ddt, boxplot(LENGTH))
```

## Task 8
### Coplot of Length vs Weight given River
```{r}
with(ddt, coplot(LENGTH~WEIGHT|RIVER, row=1))
```

### Coplot of DDT vs Weight given Species
```{r}
with(ddt, coplot(DDT~WEIGHT|SPECIES, row=1))
```

## Task 9

### Box Plot
```{r}
ggplot(ddt, aes(x = SPECIES, y= WEIGHT, fill=RIVER)) + geom_boxplot() + ggtitle('Zane Lesley')
```

### Scatter Plot
```{r}
ggplot(ddt, aes(x = RIVER, y= LENGTH, fill=SPECIES)) + geom_violin() + ggtitle('Zane Lesley')
```

### Scatter Plot
```{r}
ggplot(ddt, aes(x = WEIGHT, y= LENGTH, color=SPECIES)) + geom_point() + ggtitle('Zane Lesley')
```


