---
title: "R package"
author: "Dr Wayne Stewart"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction



Congratulations for getting this far. To complete the evidence that you have made a working R package please answer the following by using the supplied rmd  and upload the knitted html to CANVAS.

I will assume your package is called **MYPAC** and you have packaged and documented functions `myf1`, `myf2`, `myf3` and `myf4`.

Your actual package name and functions will obviously be called something else.

# First function

Please place your first function in the r chunk in the following way:

Once you are ready please remove the option `eval = FALSE`

```{r first}
LabPackages::ntickets(N=400,gamma = 0.02, p = 0.95)
```


# Second function

Please place your second function in the r chunk in the following way:

Once you are ready please remove the option `eval = FALSE`

```{r second}
LabPackages::mybin(n=10, p=0.7, iter=100)
```


# Third function

Please place your third function in the r chunk in the following way:

Once you are ready please remove the option `eval = FALSE`

```{r third}
LabPackages::myncurve(mu=10, sigma=5, a=6)
```


# Fourth function

Please place your fourth function in the r chunk in the following way:

Once you are ready please remove the option `eval = FALSE`

```{r fourth}
LabPackages::mycltp(n = 5, iter = 10000, lambda = 10)
```

# The github repository



This is the address of my GITHUB repository:
<https://github.com/ZaneLesley/FALL224753lesl0009/tree/main/LabPackages>
