---
title: 'Ass 3'
author: "Zane Lesley"
date: '`r format(Sys.Date(),format="%A, %B %d, %Y")`'
output: 
  html_document:
    df_print: paged
    fig_caption: true
    highlights: pygments
    number_sections: no
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE, fig.align='center'}
knitr::opts_chunk$set(echo = TRUE)
```

# Questions{9 / 16}

## Q 1
```{r}
probablilty <- pexp(120, rate = 1/95, lower.tail = FALSE)
probablilty
```

```{r}
phising <- read.csv("PHISHING.csv")

mean(phising$INTTIME)
sd(phising$INTTIME)
```

The theoretical mean and sd will be 1 / 95, meaning that they are close enough to the mean and sd, therefore the exponential distribution does support the idea with beta as 95.

## Q 2

### mean and variance of the maximum flood level 
$$\alpha * \beta^2 = 3 * 0.07 = \fbox{0.21} \\
\alpha * \beta^2 = 3 * 0.07^2 = 3 * 0.0049 = \fbox{0.0147} \\
\bar{x} + 3\sigma^2 \\
= 0.21 + 3(0.1213) = \fbox{0.5} \\
\bar{x} - 3\sigma^2 \\
= 0.21 - 3(0.1213) = \fbox{0.17}$$ 

### Conclusion
We expect the value to be (0.17, 0.5) so a value of 0.6 is slightly elevated.

## Q 3

### a)
$$\bar{x} = \alpha * \beta = 2 * 2 = 4 \\
\bar{x} = \alpha * \beta = 1 * 4 = 4$$

### b)
$$\sigma^2 = \alpha * \beta^2 = 2 * 2^2 = 8 \\ 
\sigma^2 = \alpha * \beta^2 = 1 * 4^2 = 16$$

### c) 
```{r}
pgamma(1, shape = 2, scale = 2)
pgamma(1, shape = 1, scale = 4)
```

Equation 2

## Q 4

### a)
```{r}
pweibull(2, shape = 2, scale = 4)
```

### b) Mean and Variance
```{r}
x = 2 * gamma(1.5)
x

sigma = 2 * sqrt((gamma(2) - (gamma(1.5)^2)))
sigma
```

### c) 
```{r}
x - 2 * sigma

x + 2 * sigma
```

### d)
```{r}
# P < 6
a = pweibull(6, shape = 2, scale = 4)

# P > 6
1 - a
```

## Q 5

### a) mean and variance

$$
\frac{\alpha}{\alpha + \beta} \\ 
\frac{2}{2 + 9} = \frac{2}{11}
$$

```{r}
a = 2
b = 11
(a * b)/((a + b)^2 * (a + b + 1))
```

### b) 40% probability
```{r}
1 - pbeta(0.4, a, b)
```

### c) 10% probability
```{r}
pbeta(0.1, a, b)
```

## Q 6

### a)
$$ \alpha - 1 = 1 \\
\Rightarrow \alpha = 2 \\
\frac{1}{8} = \frac{2}{\beta}\\
\Rightarrow \beta = 16$$

### b) mean and variance
```{r}
a = 2
b = 16

b^(1/a) * gamma((a + 1) / a)

b^(2/a) * ((gamma((a + 2)/a)) - (gamma((a + 1) / a))^2)

```

### c)
$$F(y_0) = \int_{0}^{y_0} f(y) \, dy = \int_{0}^{y_0} \frac{\alpha}{\beta} y^{\alpha - 1} e^{-y^{\alpha} / \beta} \, dy \\
\Rightarrow 1 - [1 - e^{(-6)^\alpha / \beta}] =  1 - [1 - e^-2.55] = \fbox{0.1054}$$

## Q 7

### a) 
$$P(X = x) = \frac{1}{6} \\
P(Y = y) = \frac{1}{6}   \\
\Rightarrow \frac{1}{6} * \frac{1}{6} = \frac{1}{36}$$  

### b)
$$p_1(x) = 6 * \frac{1}{36} = \frac{1}{6} \\
p_2(y) = 6 * \frac{1}{36} = \frac{1}{6}$$  

### c)
$$p_1(x|y) = p(x,y)/p_2(y) = \frac{1}{36} \div \frac{1}{6} = \frac{1}{6} \\
p_1(y|x) = \frac{1}{36} \div \frac{1}{6} = \frac{1}{6}$$  

### d)
Since all of the probabilities are the same, each dice roll will not have any impact on other dice rolls meaning they are independent.  

## Q 8

### a) 

$$
P(X = 3, Y = 1) \rightarrow \frac{1}{7} \\
P(X = 1, Y = 1) \rightarrow \frac{1}{7} \\
P(X = 3, Y = 3) \rightarrow \frac{1}{7} \\
P(X = 2, Y = 1) \rightarrow \frac{2}{7} \\
P(X = 3, Y = 2) \rightarrow \frac{2}{7}
$$

### b)

$$
P(X = 3) \rightarrow \frac{4}{7} \\
P(X = 2) \rightarrow \frac{2}{7} \\
P(X = 1) \rightarrow \frac{1}{7}  
$$

### c)

$$
P(Y = 3) \rightarrow \frac{1}{7} \\
P(Y = 2) \rightarrow \frac{2}{7} \\
P(Y = 1) \rightarrow \frac{4}{7}  
$$

### d)
$$P(Y = 1 | X = 3) = \frac{1}{7} \div \frac{4}{7} = \frac{1}{4} \\
P(Y = 2 | X = 3) = \frac{2}{7} \div \frac{4}{7} = \frac{1}{2} \\
P(Y = 3 | X = 3) = \frac{1}{7} \div \frac{4}{7} = \frac{1}{4}$$  

$$P(Y = 1, X = 1) = 1 \\
P(Y = 1, X = 2) = 1 $$  

## Q 9

### a) 

$$f_Y(y) = \int_{y}^{2y} \frac{e^{-y/10}}{10y} \, dx$$

Solving this all the way out will yield

$$f_Y(y) = \frac{e^{-y/10}}{10y} * y = \frac{e^{-y/10}}{10}\\
\Rightarrow f_Y(y) = \frac{e^{-y/10}}{10}$$

exponential distribution with parameter $\lambda = \frac{1}{10}$

### b)
```{r}
1 / (1 / 10)
```

## Q 10

### a)
$$ \int_0^{\infty} \int_0^x ce^{-x^2} \, dy \, dx = 1$$

Solving yields:

$$\frac{c}{2} * 1 = 1 \\
c = 1$$

### b)

$$f_1(x) = \int_0^x f(x, y) \, dy = \int_0^x 2e^{-x^2} \, dy$$
$$f_1(x) = 2e^{-x^2} * x$$

### c)
### Part (c): Show  conditional density Y|X is a uniform distribution over the interval

$$f_{2}(y|x) = \frac{f(x, y)}{f_1(x)}
f_{2}(y|x) = \frac{2e^{-x^2}}{2e^{-x^2} \cdot x} = \frac{1}{x}$$

conditional density $( f_{2}(y|x) = \frac{1}{x}$








































































































## Q 11

### For X's

$$P(X = -1) = \frac{1}{12} + \frac{2}{12} + \frac{1}{12} = \frac{4}{12} = \frac{1}{3} \\
P(X = 0) = \frac{2}{12} + \frac{2}{12} + \frac{2}{12} = \frac{6}{12} = \frac{1}{2} \\
P(X = 1) = \frac{1}{12} + \frac{2}{12} + \frac{1}{12} = \frac{4}{12} = \frac{1}{3}$$

### For Y's

$$P(Y = -1) = \frac{1}{12} + \frac{2}{12} + \frac{1}{12} = \frac{4}{12} = \frac{1}{3} \\
P(Y = 0) = \frac{2}{12} + \frac{2}{12} + \frac{2}{12} = \frac{6}{12} = \frac{1}{2} \\
P(Y = 1) = \frac{1}{12} + \frac{2}{12} + \frac{1}{12} = \frac{4}{12} = \frac{1}{3}$$

### E(X) and E(Y)

$$E(X) = (-1) \cdot \frac{1}{3} + 0 \cdot \frac{1}{2} + 1 \cdot \frac{1}{3} = -\frac{1}{3} + 0 + \frac{1}{3} = 0 \\
E(Y) = (-1) \cdot \frac{1}{3} + 0 \cdot \frac{1}{2} + 1 \cdot \frac{1}{3} = -\frac{1}{3} + 0 + \frac{1}{3} = 0$$

### E(XY)

$$E(XY) = \sum_x \sum_y x \cdot y \cdot P(X = x, Y = y)$$

Solving this yields $E(XY) = 0$

### Covariance

$$\text{Cov}(X, Y) = E(XY) - E(X)E(Y) = 0 - (0)(0) = 0$$

### Indepedence

Since we have a P(X) and a P(Y) (such as -1) that are different, X and Y are Independent.











































## Q 12
### a)

$$ E(Y) = \frac{a + b}{2} = \frac{1 + 3}{2} = 2 $$

### b)

$$\text{Var(Y)} = \frac{(b - a)^2}{12} = \frac{1}{3} \\
\text{Var(}\bar{Y}) = \frac{\text{Var(Y)}}{n} = \frac{\frac{1}{3}}{60} = \frac{1}{180}$$

### c)
By the CLT using n = 60, the distribution should be approximately normal.  

### d)
```{r}
mean <- 2
var <- 1/3
sd <- sqrt(var)
pnorm(2.5, mean = mean, sd = sd) - pnorm(1.5, mean = mean, sd = sd)
```

### e)
```{r}
pnorm(2.2, mean = mean, sd = sd)
```

## Q 13
### a)
```{r}
n <- 20
p <- 0.4
mean <- n * p
var <- n * p * (1 - p)
sd <- sqrt(var)

z <- (1.5 - mean) / sd
pnorm(z)
```

### b)
```{r}
z <- (10.5 - mean) / sd
1 - pnorm(z)
```

### c)
```{r}
pbinom(1, size = n, prob = p)
```
```{r}
1 - pbinom(10, size = n, prob = p)
```

## Q 14
```{r}
leadcop = read.csv("LEADCOPP.csv")

t.test(leadcop$LEAD,conf.level = 0.99)
t.test(leadcop$COPPER,conf.level = 0.99)
```

### c) Interpretation

The 99% confidence interval for the mean lead level in water specimens from Crystal Lake Manors is (−1.15,6.92). This means we are 99% confident that the true mean lead level in the water samples from Crystal Lake Manors lies between -1.15 and 6.92. We can ignore the negative numbers, as for this context, it doesn't really make sense. The same thing applies for the copper level, except the interval is (0.1519,0.6647).

### d) Meaning of 99% confident
This gives us a range that we are very confident in (in this case 99% confident) that the concentration(mean) of the copper/lead will fall within the interval calculated by the t.test.

## Q 15
```{r}
solrad = read.csv("SOLARAD.csv")

stjos <- t.test(solrad$STJOS, conf.level = 0.95)
iowa <- t.test(solrad$IOWA, conf.level = 0.95)

# Extract means
stjos_upper <- stjos$conf.int[1]
stjos_lower <- stjos$conf.int[2]
iowa_upper <- iowa$conf.int[1]
iowa_lower <- iowa$conf.int[2]

lower <- stjos_lower - iowa_lower
upper <- stjos_upper - iowa_upper

cat(lower, upper)
```

### Interpretation

Since we have a non negative and non 0 mean, we can say that one of STJOS or IOWA has a significantly higher solar irritation levels than the other (in our case STJOS is higher than IOWA). The difference is statistically significant as it shows us some factor is making the difference large.

## Q 16
```{r}
diazinon = read.csv("DIAZINON.csv")

residue_diff <- diazinon$DAY - diazinon$NIGHT

t.test(residue_diff, conf.level = 0.90)
```

### Interpretation

Since the entire interval is negative, this indicates that diazinon levels are higher at night than during the day.









