---
title: "Lab 5"
author: "Zane Lesley"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1
```{r}
setwd("C:/Users/Zanel/OneDrive/Desktop/FALL 2024 STATS/FALL224753lesl0009/Lab/Lab 5")
```

## Task 2
### The function
```{r}
mybin=function(iter=100,n=10, p=0.5){ 
# make a matrix to hold the samples
#initially filled with NA's
sam.mat=matrix(NA,nr=n,nc=iter, byrow=TRUE)
#Make a vector to hold the number of successes in each trial
succ=c()
for( i in 1:iter){
#Fill each column with a new sample
sam.mat[,i]=sample(c(1,0),n,replace=TRUE, prob=c(p,1-p))
#Calculate a statistic from the sample (this case it is the sum)
succ[i]=sum(sam.mat[,i])
}
#Make a table of successes
succ.tab=table(factor(succ,levels=0:n))
#Make a barplot of the proportions
barplot(succ.tab/(iter), col=rainbow(n+1), main="Binomial simulation", xlab="Number of successes")
succ.tab/iter
}
```

### 100 iterations
```{r}
mybin(n=10, p=0.7, iter=100)
```

### 200 iterations
```{r}
mybin(n=10, p=0.7, iter=200)
```
### 500 iterations
```{r}
mybin(n=10, p=0.7, iter=500)
```
### 1000 iterations
```{r}
mybin(n=10, p=0.7, iter=1000)
```
### 10000 iterations
```{r}
mybin(n=10, p=0.7, iter=10000)
```

### Verification
```{r}
dbin.tab = round(dbinom(0:10, size=10, prob = 0.7), 4)
names(dbin.tab) = 0:10
dbin.tab
```

You can see the values are pretty close which means or functions approximates are pretty close.

## Task 3
```{r}
sample(c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,0,0,0,0,0,0,0,0), size = 5, replace=FALSE, prob = NULL)
```
### Replace = TRUE
```{r}
sample(c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,0,0,0,0,0,0,0,0), size = 5, replace=TRUE, prob = NULL)
```

### Hyper
```{r}
myhyper=function(iter=100,N=20,r=12,n=5){
# make a matrix to hold the samples
#initially filled with NA's
sam.mat=matrix(NA,nr=n,nc=iter, byrow=TRUE)
#Make a vector to hold the number of successes over the trials
succ=c()
for( i in 1:iter){
#Fill each column with a new sample
sam.mat[,i]=sample(rep(c(1,0),c(r,N-r)),n,replace=FALSE)
#Calculate a statistic from the sample (this case it is the sum)
succ[i]=sum(sam.mat[,i])
}
#Make a table of successes
succ.tab=table(factor(succ,levels=0:n))
#Make a barplot of the proportions
barplot(succ.tab/(iter), col=rainbow(n+1), main="HYPERGEOMETRIC simulation", xlab="Number of successes")
succ.tab/iter
}
```

### 100 iterations
```{r}
myhyper(n=5, N = 20,r=12, iter=100)
```

### 200 iterations
```{r}
myhyper(n=5, N = 20,r=12, iter=200)
```
### 500 iterations
```{r}
myhyper(n=5, N = 20,r=12, iter=500)
```
### 1000 iterations
```{r}
myhyper(n=5, N = 20,r=12, iter=100)
```
### 10000 iterations
```{r}
myhyper(n=5, N = 20,r=12, iter=10000)
```

### Verification
```{r}
dhyp.tab = round(dhyper(x = 0:5, m = 12, n = 8, k = 5), 4)
names(dhyp.tab) = 0:5
dhyp.tab
```

The dhyper and the myhyper are pretty close in values, close enough to assume they are approx. correct.

## Task 4

mysample should randomly generate a sample it is strictly between 1 and 10. It will replace the numbers it picks and will run
iter times. It then visualizes the data using bar plots. It also has a time parameter which will make the function wait by time.

## The function
```{r}
mysample=function(n, iter=10,time=0.5){
for( i in 1:iter){
#make a sample
s=sample(1:10,n,replace=TRUE)
# turn the sample into a factor
sf=factor(s,levels=1:10)
#make a barplot
barplot(table(sf)/n,beside=TRUE,col=rainbow(10), 
main=paste("Example sample()", " iteration ", i, " n= ", n,sep="") ,
ylim=c(0,0.2)
)

#release the table
Sys.sleep(time)
}
}
```

## Running mysample()

Change this to be iter=1 so it will only run 1 time and the html file won't be super long.
```{r}
mysample(n = 1000, iter=1, time=1)
```

The output is bar plots which shows the distribution for each number which each is around .10.

## Task 5
```{r}
choose(8, 4)
```
### P(Y>4),Y~Pois(λ=2)
```{r}
1 - ppois(q = 3, lambda = 2)
```

### P(Y=10),Y~NegBin(p=0.4,r=3)
```{r}
choose(10 - 1, 3 - 1) * 0.4 ^ 3 * 0.6 ^ (10 - 3)
```
### P(Y≤8),Y~Bin(n=15,p=0.4)
```{r}
pbinom(q = 8, size = 15, prob=0.4)
```


## Task 6
```{r}
library(devtools)
devtools::install("C:/Users/Zanel/OneDrive/Desktop/FALL 2024 STATS/FALL224753lesl0009/LabPackages")
library(LabPackages)
tmp <- LabPackages::mybin(n=10, p=0.7, iter=100)
tmp
```






















































