---
title: "Lab 8"
author: "Zane Lesley"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1
```{r}
setwd("C:/Users/Zanel/OneDrive/Desktop/FALL 2024 STATS/FALL224753lesl0009/Lab/Lab 8")
```

## Task 2

```{r}
a = 0
b = 5

sample = runif(10, a, b)
sample

mu = (a + b) / 2
v = (b - a)^2 / 12
c(mu, v)

xbar = mean(sample)
sv = var(sample)
c(xbar, sv)
```

```{r}
myclt.mean=function(n,iter){
y=runif(n*iter,0,5) #
data=matrix(y,nr=n,nc=iter,byrow=TRUE) #B
sm=apply(data,2,mean) #C
hist(sm)
sm
}

w <-myclt.mean(n = 10, iter = 10000)
```

### Answers

In line A we create a random sample of size n * iterator with uniform distribution along (a,b). In line B we fill the matrix, then in line c we sum the columns of the matrix, in line D we call our function and get the resulting plots.

## Task 3

```{r}
mean(w)
var(w)

mycltu=function(n,iter,a=0,b=10){
## r-random sample from the uniform
y=runif(n*iter,a,b)
## Place these numbers into a matrix
## The columns will correspond to the iteration and the rows will equal the sample size n
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
## apply the function mean to the columns (2) of the matrix
## these are placed in a vector w
w=apply(data,2,mean)
## We will make a histogram of the values in w
## How high should we make y axis?
## All the values used to make a histogram are placed in param (nothing is plotted yet)
param=hist(w,plot=FALSE)
## Since the histogram will be a density plot we will find the max density

ymax=max(param$density)
## To be on the safe side we will add 10% more to this
ymax=1.1*ymax
## Now we can make the histogram
hist(w,freq=FALSE,  ylim=c(0,ymax), main=paste("Histogram of sample mean",
"\n", "sample size= ",n,sep=""),xlab="Sample mean")
## add a density curve made from the sample distribution
lines(density(w),col="Blue",lwd=3) # add a density plot
## Add a theoretical normal curve 
curve(dnorm(x,mean=(a+b)/2,sd=(b-a)/(sqrt(12*n))),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
## Add the density from which the samples were taken
curve(dunif(x,a,b),add=TRUE,lwd=4)

}
mycltu(n=20,iter=100000)

for (n in c(1,2,3,5,10,30)){
  mycltu(n = n, iter = 10000, a = 0, b = 10)
}
```

### Conclusion

As n goes up in value, we get closer and closer to a normal distribution.

## Task 4
```{r}
mycltb=function(n,iter,p=0.5,...){

## r-random sample from the Binomial
y=rbinom(n*iter,size=n,prob=p)
## Place these numbers into a matrix
## The columns will correspond to the iteration and the rows will equal the sample size n
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
## apply the function mean to the columns (2) of the matrix
## these are placed in a vector w
w=apply(data,2,mean)
## We will make a histogram of the values in w
## How high should we make y axis?
## All the values used to make a histogram are placed in param (nothing is plotted yet)
param=hist(w,plot=FALSE)
## Since the histogram will be a density plot we will find the max density

ymax=max(param$density)
## To be on the safe side we will add 10% more to this
ymax=1.1*ymax

## Now we can make the histogram
## freq=FALSE means take a density
hist(w,freq=FALSE,  ylim=c(0,ymax),
main=paste("Histogram of sample mean","\n", "sample size= ",n,sep=""),
xlab="Sample mean",...)
## add a density curve made from the sample distribution
#lines(density(w),col="Blue",lwd=3) # add a density plot
## Add a theoretical normal curve 
curve(dnorm(x,mean=n*p,sd=sqrt(p*(1-p))),add=TRUE,col="Red",lty=2,lwd=3) 

}

cltb.p <- function(p = 0.5, ...){
  for(n in c(4,5,10,20)){
    mycltb(n = n, iter = 10000, p = p,...)
  }
}
cltb.p(0.3)
```

### Conclusion

As with the previous case, as the sample mean goes up in samples, ie we get more values and a larger distrubution, it becomes more and more normal.

## Task 5
```{r}
mycltp=function(n,iter,lambda=10,...){

## r-random sample from the Poisson
y=rpois(n*iter,lambda=lambda)
## Place these numbers into a matrix
## The columns will correspond to the iteration and the rows will equal the sample size n
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
## apply the function mean to the columns (2) of the matrix
## these are placed in a vector w
w=apply(data,2,mean)
## We will make a histogram of the values in w
## How high should we make y axis?
## All the values used to make a histogram are placed in param (nothing is plotted yet)
param=hist(w,plot=FALSE)
## Since the histogram will be a density plot we will find the max density

ymax=max(param$density)
## To be on the safe side we will add 10% more to this
ymax=1.1*ymax

## Make a suitable layout for graphing
layout(matrix(c(1,1,2,3),nr=2,nc=2, byrow=TRUE))

## Now we can make the histogram
hist(w,freq=FALSE,  ylim=c(0,ymax), col=rainbow(max(w)),
main=paste("Histogram of sample mean","\n", "sample size= ",n," iter=",iter," lambda=",lambda,sep=""),
xlab="Sample mean",...)
## add a density curve made from the sample distribution
#lines(density(w),col="Blue",lwd=3) # add a density plot
## Add a theoretical normal curve 
curve(dnorm(x,mean=lambda,sd=sqrt(lambda/n)),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve

# Now make a new plot
# Since y is discrete we should use a barplot
barplot(table(y)/(n*iter),col=rainbow(max(y)), main="Barplot of sampled y", ylab ="Rel. Freq",xlab="y" )
x=0:max(y)
plot(x,dpois(x,lambda=lambda),type="h",lwd=5,col=rainbow(max(y)),
main="Probability function for Poisson", ylab="Probability",xlab="y")
}


mycltp.lambda <- function(lambda = 10, iter = 10000, ...){
  for (n in c(2, 3, 5, 10, 20)){
    mycltp(n = n, iter = iter, lambda = lambda, ...)
  }
}
mycltp.lambda(10)
```

## Task 6

```{r}
library(devtools)
devtools::install("C:/Users/Zanel/OneDrive/Desktop/FALL 2024 STATS/FALL224753lesl0009/LabPackages")
library(LabPackages)
LabPackages::mycltp(n = 5, iter = 10000, lambda = 10)
```























